---
output:
  html_document: default
  pdf_document: default
---
Exploration of Red Wine Quality by Kyungwon Chun
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
#install.packages('psych')
#install.packages('dplyr')

library(ggplot2)
library(GGally)
library(psych)
library(dplyr)
library(memisc)
library(RColorBrewer)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wqr <- read.csv('wineQualityReds.csv')
str(wqr)
```

This tidy dataset contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating from 0 (very bad) to 10 (very excellent).

# Univariate Plots Section

```{r echo=FALSE}
summary(wqr)
```

```{r echo=FALSE}
ggplot(aes(fixed.acidity), data=wqr) +
  geom_histogram(binwidth=0.5)
```

```{r echo=FALSE}
summary(wqr$fixed.acidity)
```

The distribution of fixed acidity is positive skewed. Most of the wines have 
fixed acidity between 7.10 and 9.20.

```{r echo=FALSE}
ggplot(aes(volatile.acidity), data=wqr) +
  geom_histogram(binwidth=0.02)
```

```{r echo=FALSE}
summary(wqr$volatile.acidity)
```

The valatile acidity shows a bimodal distribution and positive skewness. Most of
the wines have volatile acidity between 0.39 and 0.64.

```{r echo=FALSE}
wqr$total.acidity <- wqr$volatile.acidity + wqr$fixed.acidity
ggplot(aes(total.acidity), data=wqr) + geom_histogram(binwidth=0.1)
```

```{r echo=FALSE}
summary(wqr$total.acidity)
```

Total acidity is composed of fixed and volatile acidity. The distribution of
total acidity is positive skewed with median at 8.445.

```{r echo=FALSE}
ggplot(aes(citric.acid), data=wqr) +
  geom_histogram(binwidth=0.05)
```

```{r echo=FALSE}
ggplot(aes(residual.sugar), data=wqr) +
  geom_histogram(binwidth=0.2)
```

The residual sugar shows left-biased and long-tailed distribution.

```{r echo=FALSE}
ggplot(aes(chlorides), data=wqr) +
  geom_histogram(binwidth=0.01)
```

The chlorides show left-biased and long-tailed distribution.

```{r echo=FALSE}
ggplot(aes(free.sulfur.dioxide), data=wqr) +
  geom_histogram(binwidth=1)
```

```{r echo=FALSE}
ggplot(aes(total.sulfur.dioxide), data=wqr) +
  geom_histogram(binwidth=5)
```

The total sulfur dioxide has some outliers.

```{r echo=FALSE}
ggplot(aes(density), data=wqr) +
  geom_histogram(binwidth=0.0005)
```

```{r echo=FALSE}
summary(wqr$density)
```

Most of the wines have a density between 0.9956 and 0.9978.

```{r echo=FALSE}
ggplot(aes(pH), data=wqr) +
  geom_histogram(binwidth=0.05)
```

```{r echo=FALSE}
summary(wqr$pH)
```

Most of the wines have pH between 3.210 and 3.400.

```{r echo=FALSE}
ggplot(aes(sulphates), data=wqr) +
  geom_histogram(binwidth=0.05)
```

```{r echo=FALSE}
ggplot(aes(alcohol), data=wqr) +
  geom_histogram(binwidth=0.1)
```

```{r echo=FALSE}
ggplot(aes(as.factor(quality)), data=wqr) +
  geom_bar()
```

```{r echo=FALSE}
summary(wqr$quality)
```

Most of the wines have 5 or 6 in quality.

# Univariate Analysis

### What is the structure of your dataset?

There are 15,999 red wines in the dataset with 13 features (X, fixed acidity,
volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide,
total sulfur dioxide, density, pH, sulphates, alcohol, quality). X identifies 
the wines, and quality represents that how good the wine. The X and quality are
unordered and ordered factor variables, but I treated them as numerical 
variables for convenience. All other features represent chemical properties of 
wine.

Other observations: 

* Wines with quality 5 or 6 are most common.
* The median wine quality is 6.
* Most wines have a quality of 5 or better.
* About 75% of wines have a quality of 6 or worse.
* The worst and best quality in the data set is 3 and 8, respectively.
 
### What is/are the main feature(s) of interest in your dataset?

The main feature in the data set is quality. I'd like to determine which 
features are best for predicting the wine quality. I suspect quality and some 
combination of the other variables can be used to build a predictive model for 
wine quality.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The primary wine characteristics are sweetness, acidity, tannin, alcohol, and 
body. Residual sugar, fixed and volatile acidity, alcohol, and density determine
those characteristics. I guess that these variables are mainly related to the 
wine quality.

### Did you create any new variables from existing variables in the dataset?

I created a variable for the total acidity using the volatile and the 
fixed acids.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Volatile acidity shows a bimodal distribution.

# Bivariate Plots Section

```{r echo=FALSE}
round(cor(wqr[!names(wqr) %in% c("X")]), 2)
```

```{r echo=FALSE}
pairs.panels(wqr[names(wqr) %in% c('pH', 'fixed.acidity', 'citric.acid', 
                                   'volatile.acidity', 'alcohol', 'density', 
                                   'quality')])
```

The fixed acidity and volatile acidity has strong positive and negative 
correlations with citric acid.

The pH has a strong negative correlation with fixed acidity, citric acid, but 
does not with volatile acidity.

The fixed acidity and alcohol have significant positive and negative 
correlations with density, respectively.

Most of the variables do not seem to have strong correlations with quality, 
but alcohol and volatile acidity have moderate positive and negative 
correlation with quality, respectively.

```{r echo=FALSE}
ggplot(wqr, aes(x=fixed.acidity, y=pH)) +
  geom_point(alpha = 0.3, size = 2) +
  geom_smooth(method='lm')
```

The strongest correlation in this data set appears between fixed acidity and 
pH. High acidity means low pH, and the graph coincides with this fact.  

```{r echo=FALSE}
ggplot(wqr, aes(x=fixed.acidity, y=citric.acid)) +
  geom_point(alpha = 0.3, size = 2) +
  stat_smooth(method='lm')
```

Citric acid is ne of the main component of fixed acidity. Therefore the two 
variable has a strong positive correlation.

```{r echo=FALSE}
ggplot(wqr, aes(x=fixed.acidity, y=density)) +
  geom_point(alpha = 0.3, size = 2) +
  stat_smooth(method='lm')
```

The fixed acidity has a strong positive correlation with density, too.

```{r echo=FALSE}
ggplot(wqr, aes(x=citric.acid, y=volatile.acidity)) +
  geom_point(alpha = 0.3, size = 2) +
  stat_smooth(method='lm')
```

Yeast in wine convert citric acid to acetic acid, most of the volatile acid. 
Therefore, volatile acidity and citric acid is in a reverse relation.

```{r echo=FALSE}
ggplot(wqr, aes(x=citric.acid, y=pH)) +
  geom_point(alpha = 0.3, size = 2) +
  stat_smooth(method='lm')
```

The citric acid has moderate negative correlations with volatile acidity
and pH.

```{r echo=FALSE}
ggplot(wqr, aes(x=alcohol, y=density)) +
  geom_point(alpha = 0.3, size = 2) +
  stat_smooth(method='lm')
```

The alcohol and density also show moderate negative correlation.

```{r echo=FALSE}
ggplot(wqr, aes(x=quality, y=volatile.acidity)) +
  geom_jitter(alpha=0.3, size=2) +
  stat_smooth(method=lm)
```

Quality of wine tends to increase as volatile acidity decreases, because the
main component of volatile acid is acetic acid which causes an unpleasant 
vinegar taste.

```{r echo=FALSE}
summary(lm(quality ~ volatile.acidity, wqr))
```

Based on the value of R-squared, volatile acidity contributes only about 
15.2% of the Wine quality.

```{r echo=FALSE}
ggplot(wqr, aes(x=quality, y=alcohol)) +
  geom_jitter(alpha = 0.3, size = 2) +
  stat_smooth(method='lm')
```

```{r echo=FALSE}
summary(lm(quality ~ I(sqrt(alcohol)), wqr))
```

Based on the value of R-squared, alcohol contributes to the wine quality only 
about 15.2%.

```{r echo=FALSE}
ggplot(subset(wqr, residual.sugar < quantile(residual.sugar, 0.99)), 
       aes(x=factor(quality), y=residual.sugar)) + 
  geom_jitter(alpha=0.5) +
  geom_boxplot(alpha=0.5, color=I("Red"))
```

Residual sugar determines the sweetness of the wine. Most of the wine maintain 
an certain level of sweetness.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The quality correlates with alcohol and volatile acidity.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Citric acid is one of the main components of fixed acidity. As a result, they 
have a strong positive correlation. 

High fixed acidity causes low pH. Therefore, fixed acidity and citric acid 
negatively correlates with the pH.

Wine with more volatile acidity tends to have less citric acid.

Wine with more fixed acidity tends to denser. By the way, A wine with more 
alcohol tends to less dense.

### What was the strongest relationship you found?

The fixed acidity is positively and strongly correlated with citric acid and 
density. The citric acid may substitute for fixed acidity and density with even
better estimation of wine quality.

# Multivariate Plots Section

```{r echo=FALSE}
grid.arrange(
ggplot(aes(x=sulphates, y=volatile.acidity, color=factor(quality)), data=wqr) +
  geom_point(alpha=0.5) +
  scale_colour_brewer(type="qual"),
ggplot(aes(x=log10(sulphates), y=volatile.acidity, color=factor(quality)), 
       data=wqr) +
  geom_point(alpha=0.5) +
  scale_colour_brewer(type="qual")
)
```

```{r echo=TRUE}
c(cor(wqr$volatile.acidity, wqr$sulphates), 
  cor(wqr$volatile.acidity, log10(wqr$sulphates)))
```

Transformation of sulphates to log10(sulphates) increase the correlation between
sulphates and volatile acidity.

```{r echo=FALSE}
grid.arrange(
ggplot(aes(x=pH, y=alcohol, color=factor(quality)), data=wqr) +
  geom_point(alpha=0.5) +
  scale_colour_brewer(type="qual"),
ggplot(aes(x=pH^7, y=alcohol, color=factor(quality)), data=wqr) +
  geom_point(alpha=0.5) +
  scale_colour_brewer(type="qual")
)
```

```{r echo=TRUE}
c(cor(wqr$alcohol, wqr$pH), cor(wqr$alcohol, wqr$pH^7))
```

Transformation of pH to pH^7 increases the correlation between pH and alcohol 
little bit. As shown below, this leads the increase of our model accuracy little
bit.

```{r echo=FALSE}
m1 <- lm(quality ~ alcohol, data=wqr)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + sulphates)
m4 <- update(m3, ~ . + chlorides)
m5 <- update(m4, ~ . + total.sulfur.dioxide)
m6 <- update(m5, ~ . + pH)
m7 <- update(m6, ~ . + citric.acid)
mtable(m1, m2, m3, m4, m5, m6, m7, sdigits=3)
```

The first trial of linear model accounts for 35.7% of the variance. The 
variables with less significance were removed.

```{r echo=FALSE}
m1 <- lm(quality ~ alcohol, data=wqr)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + I(log10(sulphates)))
m4 <- update(m3, ~ . + chlorides)
m5 <- update(m4, ~ . + total.sulfur.dioxide)
m6 <- update(m5, ~ . + I(pH^7))
m7 <- update(m6, ~ . + citric.acid)
mtable(m1, m2, m3, m4, m5, m6, m7, sdigits=3)
```

The variables in this linear model can account for 37.0% of the variance in the 
quality of the wine. By using log10(sulphates) and pH^7, we could improve the 
result compared to 35.7% without transformation.

```{r echo=FALSE}
summary(m7)
```

```{r echo=FALSE}
par(mfrow = c(2, 2))
plot(m7)
```

```{r echo=FALSE}
qplot(residuals(m7), fill=I("Orange"), color=I("Black"))
```

```{r echo=FALSE}
wqr$fitted <- fitted(m7)

ggplot(aes(round(fitted)), data=wqr) +
  facet_wrap(~quality) +
  geom_histogram() +
  labs(title="Facet wrap by quality")
```

```{r echo=FALSE}
wqr$fitted <- fitted(m7)

ggplot(aes(x=factor(quality), y=fitted), data=wqr) +
  geom_jitter(alpha=0.3) +
  geom_boxplot(alpha=0.3, color=I("Red")) +
  labs(y="fitted quality")
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Transformation of sulphate and pH increases the correlations with other 
variables. These transformations give clue to make a better linear model.

### Were there any interesting or surprising interactions between features?

High alcohol and low volatile acidity contents seem to produce better wines.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I created a couple of linear models. Though the confidence level of the model 
could be increased a bit by transforming a couple of variables, the final model
still is not satisfactory. This can be due to the fact that our dataset contains
a small number of observations. Furthermore, most of the observations are from
middle-classed wines. This makes it difficult that the model predict the edge 
cases. Maybe a more supplement dataset with more edge cases would help to 
predict the accurate quality of wines.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(wqr, aes(x=factor(quality), y=alcohol)) + 
  geom_jitter(alpha=0.5) +
  geom_boxplot(alpha=0.5, color=I("Red")) +
  labs(x='Quality', y='Alcohol (% by volume)', 
       title='Influence of alcohol on wine quality')
```

### Description One

Alcohol percentage plays a primary role in determining the quality of wines. The
higher the alcohol percentage, the better the wine quality. But previously from 
our linear model test, R-Squared value tells that alcohol alone contributes only
about 22% in the variance of the wine quality. So alcohol is not the only factor
which is responsible for the improvement in wine quality.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(wqr, aes(x=factor(quality), y=volatile.acidity)) + 
  geom_jitter(alpha=0.5) +
  geom_boxplot(alpha=0.5, color=I("Red")) +
  labs(x='Quality', y='Acetic acid (g / dm^3)', 
       title='Influence of volatile acidity on wine quality')
```

### Description Two

The volatile acidity has a negative relation with wine quality, though it is 
weaker than that of alcohol. It seems that the main component of volatile acid 
is an acetic acid which causes the unpleasant vinegar taste.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(round(fitted)), data=wqr) +
  facet_wrap(~quality, scales="free") +
  geom_histogram() +
  labs(x="Fitted quality", title="Facet wrap by quality")
```

### Description Three

We can see that the model fails to predict the good and bad quality wines. This 
is evident from the fact that most data sets contain 'average' quality wine and 
there are insufficient observations in the extreme range. The R-squared value of
our model can only account for about 37.1% observations.

------

# Reflection

The data analyzed in this project contains the chemical properties and quality 
information for 15,999 red wines. Based on the statistics of the chemical
properties of wines, I tried to establish a model to predict the quality of each
red wine.

Some of the ingredients showed strong correlations with others, and their 
relationship could be explained chemically. Alcohol and volatile acidity were 
directly correlated with the quality of the wine, and these characteristics 
helped to establish a quality prediction model. Sulphate and pH were able to 
increase the correlation with quality by using variable transformations, and 
this attempt helped to raise the quality prediction accuracy to some extent.

The resultant model had a low prediction success rate because most of the wine 
contained in the data had a quality of 5 or 6 and the number of samples for 
other qualities was not sufficient.

This project used only red wine dataset, but the white wine dataset contains 
three times more observable than red wine dataset. If there are common
constituents among the two types of wines, and if they exhibit the same tendency
about quality, it would be possible to supplement insufficient observables.

------

# References

* https://en.wikipedia.org/wiki/Acids_in_wine
* https://winefolly.com/review/wine-characteristics/
